Identifying Values to Express Emotions with a Non-Anthropomorphic Platform

### 2017-04-23 JournalIJSR/Parts/Emotion_Projection.tex
Last version of the article. This version includes most of reviewers´ comments, such as new references, explanation of parts that were not well explained before.

Although it is not possible to apply a direct mapping from human studies to robots~\cite{Saerbeck2007,Canamero2010}, human studies provide guidelines about possible features and values that could be used to generate emotional motion.

@@

Emotion plays an important role in human-human interaction, and can be expressed through diverse channels such as body gestures and poses, body movements, face expressions. Given the complex structure of human face, where more than 43 muscles act, face configurations are considered as a primary channel to express emotion. As a consequence, many works have focused on facial expression, mainly, but not only, influenced by the work done by Ekman~\cite{Ekman2004}. However, the role that body plays in emotion projection was recognized, too, and started to be studied~\cite{Gelder2008,Wallboot1998}. Nevertheless, the amount of works related to body expression of emotions is still small compared to studies about facial expression.

Analysing some of the few works that have studied human body expressiveness, it is possible to recognize two different methodologies to create the data base of movements to be assessed during the experiments. Applying a first methodology, human actors (either professionals, or amateurs) are asked to walk straight from point A to point B conveying specific emotions~\cite{Dael2012,Meijer1989,Wallboot1998}. Each trial is recorded and later shown to each subjects that have to classify all the sequences. The second methodology uses virtual agents~\cite{Roether2009,Venture2014} to generate the very same set up for the experiments; the agent's movements are generated from the data recorded from human actors.

The work done by Wallboot~\cite{Wallboot1998} has been used as a reference by others. It addressed the question: \textit{Are specific body postures indicative of emotion or are they only used to indicate the intensity of the emotion?} To answer this question, he recorded 224 videos for joy, happiness, sadness, despair, fear, terror, coldness anger, hot anger, disgust, contempt, shame, guilt, pride, and boredom, and showed them to the subjects. His results reaffirm that movement and body postures are indicative of intensity for some emotions. But at the same time, these two characteristics seem to be enough to identify other emotions.

@@

The main drawback of all these projects is the use of video recorded sequences, which misses the impact related to a complete physical experience. For instance, it is clearly different facing an angry robot really rushing against us, or looking at a video where this happens.

\subsection{Robotic Studies}
%%%%%%%%%%%%%%%%%%%%%%

As a direct consequence of the abundance of works in face elicitation in humans, most of the works done in Human-Robot Interaction (HRI) have focused also on faces. One of the most well-known expressive robots is Kismet~\cite{Breazeal2002}, a robotic face able to interact with people and to show emotions. The face had enough degrees of freedom to portray the basic emotions suggested by Ekman~\cite{Ekman2004} (happiness, surprise, anger, disgust, fear, and sadness), plus interest. 
Despite the complex system behind Kismet, the emotion's projection evaluation was done using videos with a very limited number of participants. Similar approach was followed by Li and Chignell~\cite{Li2011}, who used videos of a teddy bear robot to study the contribution of arms and head movement to express emotions. In same direction Destephe and collaborators~\cite{Destephe2013b,Destephe2013} studied the attribution of emotion to a robot's gait using a virtual representation of the platform WABIAN-2R. More recently Knight and Simmons~\cite{knight2016} used Keepon and NAO platforms to study the possibility to project inner states with just head movements. Although use of videos has the advantage to cover a major number of participants, the lack of interaction with the real platform make these works lose the impact that a robot could generate on the participants.

%%%%%%%%%%%%%%%%%%%%%%%%
The use of real platforms to study emotion projection could be dived in two lines: using anthropomorphic and non-anthropomorphic platforms. In the first case, these works are characterized by the use of cue positions to project desire emotions~\cite{NAO2013}. In some cases, special attention has been taken to determine head's angle and arms position contribution in specific emotions~\cite{Brown2014}. 
Nevertheless, current humanoid platforms cannot generate smooth gaits, which limit the study of body movements. To overcome this limitation, some researchers have reduce the human appearance (e.g. eliminating limps and facial expressions)to increase platform mobility and study new mechanisms to project emotions~\cite{Arras2012}. Pushing forward the reduction on anthropomorphic features, Saerbeck and Christoph used a Roomba platform to study the contribution of curvature in a trajectory to conceive emotional states~\cite{Saerbeck2010}. Similarly Lourens and Barakova~\cite{BarakovaL10} implemented a set of behaviors to determine the emotion  perceived from diverse movements, which were selected from the  work done by Camurri et al.~\cite{pop00002}. Continuing with her research on how robotics' behaviours are interpreted by people, Barakova and collaborators~\cite{Barakova2013} created a closet in which lights could be manipulated to convey pre-defined behaviours. The robot's behaviours were defined using the Interpersonal Behaviour Circle (ICB)~\cite{Leary57}, which is based on two dimensions (dominance-submission and hate-love). Their findings suggest that electronic systems can elicit a type of reactions different from the one expected by theories of interpersonal communication.

Other approaches have tried to get a better understanding of the contribution of diverse features to express emotions through movement. For example, Suk and collaborators payed a particular attention to speed, smoothness, granularity of movement path and volume of a non-bioinspired object~\cite{NAM2014}. Their results suggest that arousal increases as speed increases and that there is not any clear tendency for smoothness. On the other hand, granularity is positively correlated with pleasure and arousal, while volume is negatively correlated with pleasure and positively correlated with arousal. Alike, Tan and collaborators~\cite{Tan2016} have studied the contribution of velocity, fluidity, direction and orientation of a small box. Their results suggest that direction is directly correlated with dominance, but that fluidity does not influence the perception. While flat orientation is related to positive valence, leaning position are related with negative valance. Finally the velocity is correlated with valence, arousal and dominance.

>>>>

It is not possible to apply a direct mapping from human studies to robots~\cite{Saerbeck2007,Canamero2010}. Nevertheless works done in human studies provide guidelines about possible features and values that could be used to generate emotional motion.

@@

Emotion plays an important role in human-human interaction and can be expressed through diverse channels such as body gestures and poses, body movements, face expressions. The human face is a complex structure that embraces more than 43 muscles act. Hence, the face has been considered as a primary channel to express emotion. As a consequence, many works have focused on facial expression, mainly but not only, influenced by the work done by Ekman~\cite{Ekman2004}. However, the role that body plays in emotion projection has been recognized and it has been started to be studied~\cite{Gelder2008,Wallboot1998}. Nevertheless, the amount of works related to body expression of emotions is still small compared to studies about facial expression.

Analysing some of the few works that have studied human body expressiveness, it is possible to recognize two different methodologies to create the data base of movements to be assessed during the experiments. The first methodology uses human actors (either professionals or amateurs) to walk straight from point A to point B conveying specific emotions~\cite{Dael2012,Meijer1989,Wallboot1998}. Each trial is recorded and later shown to each subjects that have to classify all the sequences. The second methodology uses virtual agents~\cite{Roether2009,Venture2014} to generate the very same set up for the experiments; the agent's movements are generated from the data recorded from human actors.

The work done by Wallboot~\cite{Wallboot1998} has been used as a reference by other researches. It addressed the question: \textit{Are specific body postures indicative of emotion or are they only used to indicate the intensity of the emotion?} To answer this question, he recorded 224 videos for joy, happiness, sadness, despair, fear, terror, coldness anger, hot anger, disgust, contempt, shame, guilt, pride, and boredom, and showed them to the subjects. His results reaffirm that movement and body postures are indicative of intensity for some emotions. But at the same time, these two characteristics seem to be enough to identify other emotions.

@@

There are two main drawbacks of all these projects that use video recorded sequences. First, they miss the impact related to a complete physical experience. For instance, it is clearly different facing an angry robot really rushing against us, or looking at a video where this happens. Second, each actor has his/her own way to represent a given emotion~\cite{Russell2003,Gunes2011}. Actors use diverse techniques to create believable representations of specific emotions. However, this does not ensure that all actor convey emotions in same way. Therefore, several records are done and the ones with the highest agreement are selected. This brings the technical question on how to interpret the significance of agreement obtained~\cite{Russell2003}.

\subsection{Robotic Studies}
%%%%%%%%%%%%%%%%%%%%%%
As a direct consequence of the abundance of works in face elicitation in humans, most of the works done in Human-Robot Interaction (HRI) have focused also on faces. One of the most well-known expressive robots is Kismet~\cite{Breazeal2002}, a robotic face able to interact with people and to show emotions. The face had enough degrees of freedom to portray the basic emotions suggested by Ekman~\cite{Ekman2004} (happiness, surprise, anger, disgust, fear, and sadness), plus interest. 
Despite the complex system behind Kismet, the emotion's projection evaluation was done using videos with a very limited number of participants. Similar approach was followed by Li and Chignell~\cite{Li2011}, who used videos of a teddy bear robot to study the contribution of arms and head movement to express emotions. In same direction Destephe and collaborators~\cite{Destephe2013b,Destephe2013} studied the attribution of emotion to a robot's gait using a virtual representation of the platform WABIAN-2R. More recently Knight and Simmons~\cite{knight2016} used Keepon and NAO platforms to study the possibility to project inner states with just head movements. Although use of videos has the advantage to cover a major number of participants, the lack of interaction with the real platform make these works lose the impact that a robot could generate on the participants.

%%%%%%%%%%%%%%%%%%%%%%%%
The use of real platforms to study emotion projection could be dived in two lines: using anthropomorphic and non-anthropomorphic platforms. In the first case, these works are characterized by the use of cue positions to project desire emotions~\cite{NAO2013}. In some cases, special attention has been taken to determine head's angle and arms position contribution in specific emotions~\cite{Brown2014}. 
Nevertheless, current humanoid platforms cannot generate smooth gaits, which limit the study of body movements. For example Karg et al.~\cite{Karg2010} tried to map actors gaits on an hexadop robot to understand the contribution of step length, height and velocity. Their results suggest that using this approach is not suitable because the difference between different gaits are not recognizable. So they suggest that movements must be exagereted.  Some researchers have reduced the human appearance (e.g. eliminating limps and facial expressions)to increase platform mobility and study new mechanisms to project emotions~\cite{Arras2012}. Pushing forward the reduction on anthropomorphic features, Saerbeck and Christoph used a Roomba platform to study the contribution of curvature in a trajectory to conceive emotional states~\cite{Saerbeck2010}. Similarly Lourens and Barakova~\cite{BarakovaL10} implemented a set of behaviors to determine the emotion  perceived from diverse movements, which were selected from the  work done by Camurri et al.~\cite{pop00002}. Continuing with her research on how robotics' behaviours are interpreted by people, Barakova and collaborators~\cite{Barakova2013} created a closet in which lights could be manipulated to convey pre-defined behaviours. The robot's behaviours were defined using the Interpersonal Behaviour Circle (ICB)~\cite{Leary57}, which is based on two dimensions (dominance-submission and hate-love). Their findings suggest that electronic systems can elicit a type of reactions different from the one expected by theories of interpersonal communication.

Other approaches have tried to get a better understanding of the contribution of diverse features to express emotions through movement. For example, Suk and collaborators payed a particular attention to speed, smoothness, granularity of movement path and volume of a non-bioinspired object~\cite{NAM2014}. They used the self-assessment (SAM)~\cite{Lang2008} method to evaluate participants. Their results suggest that arousal increases as speed increases and that there is not any clear tendency for smoothness. On the other hand, granularity is positively correlated with pleasure and arousal, while volume is negatively correlated with pleasure and positively correlated with arousal. Alike, Tan and collaborators~\cite{Tan2016} have studied the contribution of velocity, fluidity, direction and orientation of a small box. They evaluated participants' perception using SAM. Their results suggest that direction is directly correlated with dominance, but that fluidity does not influence the perception. While flat orientation is related to positive valence, leaning position are related with negative valance. Finally the velocity is correlated with valence, arousal and dominance. Using an hexapod robot, Karg et al. studied

### 2017-04-23 JournalIJSR/Parts/ExperimentDesign.tex
Last version of the article. This version includes most of reviewers´ comments, such as new references, explanation of parts that were not well explained before.

The experiment was designed to get a better understanding about the contribution linear and angular velocity, oscillation angle, direction, and orientation can give to the expression of \textit{Happiness}, \textit{Anger}, \textit{Sadness}, and \textit{Fear}, which are four out of the six emotions considered by Ekman~\cite{Ekman2004} as basic emotions, which are characterized to have short duration, cultural universal~\cite{Ekman94}, derive from evolution~\cite{Buck99}, and could influence cognition and action~\cite{Campos2004}. Moreover, these four emotions here selected are overlap with emotion schemas~\cite{Izard2007}.
%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%

@@

\begin{table}[htb]
\centering
\caption{Values for each of the independent variables in the experiment.}
\begin{tabular}{|c|c|c|c|c|}
\hline
\backslashbox{Variable}{Possibilities} & First & Second & Third & Fourth\\
\hline   
Angular Velocity ($rad/sec$)& $0$ & $1$ & $2$ & $3$\\
\hline
Oscillation Angle ($rad$)& $0$ & $0.087$ & $0.175$ & $0.349$\\
\hline
Linear Velocity ($mm/sec$) & $0$ & $200$ & $500$ & $900$\\
\hline
Direction ($rad$)&$0$&$\pi$&$\frac{-\pi}{2}$& \\
\hline
Orientation ($rad$) & $0$ & $\pi$ & & \\
\hline 
\multicolumn{5}{c}{}
\end{tabular} 
\label{table:variables_values}
\end{table}

@@

\begin{itemize}
	\item \textbf{Emotion} is the feeling perceived by the participants from the robot's movement. From previous experiences~\cite{angel2014}
	, it was decided to ask the participants to select an emotion name in a list including the four emotions intended to be expressed, two mental states that could be misinterpreted from these emotions, and the option of ``other''. 
	The two  states of mind included as confounding terms are tenderness and excitement, which correspond to low and high arousal states.

@@

	\item The robot was shown to the participant and the experiment procedure explained.

	\item An example of the questionnaire was presented and a sample movement of the robot shown.

@@

The order of the options listed in the questionnaire changed for each question to prevent any kind of bias. Figure~\ref{fig:questionnaire_example} shows an example of the questionnaire used.

>>>>

The experiment was designed to get a better understanding about the contribution given by linear and angular velocity, oscillation angle, direction, and orientation to the expression of \textit{Happiness}, \textit{Anger}, \textit{Sadness}, and \textit{Fear}. These four emotions were selected because they are considered as natural kinds and emotion schemas~\cite{Izard2007}. Moreover they are derive from evolution~\cite{Buck99} and could influence cognition and action~\cite{Campos2004}, which could be considered in congintive architecures. In addition, \textit{Anger} and \textit{Happiness} have been confused in previous studies~\cite{Angel2016}. Hence, given a better understanding on how to express these emotions could give researchers a better idea how to avoid misinterpretations that could be critical in a real context.

%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%

@@

\begin{table}[htb]
\centering
\caption{Values for each of the independent variables in the experiment.}
\begin{tabular}{|c|c|c|c|c|}
\hline
\backslashbox{Variable}{Possibilities} & \rotatebox{-90}{First} & \rotatebox{-90}{Second} & \rotatebox{-90}{Third} & \rotatebox{-90}{Fourth}\\
\hline   
Angular Velocity ($rad/sec$)& $0$ & $1$ & $2$ & $3$\\
\hline
Oscillation Angle ($rad$)& $0$ & $0.087$ & $0.175$ & $0.349$\\
\hline
Linear Velocity ($mm/sec$) & $0$ & $200$ & $500$ & $900$\\
\hline
Direction ($rad$)&$0$&$\pi$&$\frac{-\pi}{2}$& \\
\hline
Orientation ($rad$) & $0$ & $\pi$ & & \\
\hline 
\multicolumn{5}{c}{}
\end{tabular} 
\label{table:variables_values}
\end{table}

@@

\begin{itemize}
	\item \textbf{Emotion} is the feeling perceived by the participants from the robot's movement. From previous experiences~\cite{angel2014}
	, it was decided to ask the participants to select an emotion name in a list including the four emotions intended to be expressed, two mental states that could be misinterpreted from these emotions, and the option of ``other''.
	The two  states of mind included as confounding terms are tenderness and excitement, which correspond to low and high arousal states. This was decided because from previous works done in robotics, it was reported that body movement is related to arousal. As it was already mentioned, some psychologist suggest that diverse theoretical emotions could overlap, so combining discrete and dimensional theories is possible.

@@

	\item The robot was shown to the participant and it was explained its characteristics and layout. Then the experiment procedure explained. If the participants had any questions, they were answered.

	\item An example of the questionnaire was presented and a sample movement of the robot shown. The participant is remember where the front of the robot is.

@@

The order of the options listed in the questionnaire changed for each question to prevent any kind of bias. Figure~\ref{fig:questionnaire_example} shows an example of the questionnaire used. As it could be observed, the questionnaire allows the participant to introduce more than one emotion.

### 2017-04-23 JournalIJSR/Parts/Introduction.tex
Last version of the article. This version includes most of reviewers´ comments, such as new references, explanation of parts that were not well explained before.

Social environments are intricate spaces where people behave in ways that lead to their acceptance in social groups. Similarly, studies in social robotics have shown that robots acceptance increases when robots project a high social presence~\cite{Heerink08}. The straightest way to fulfill is through imitation of humans' characteristics, such us body form, behaviors and social characteristics. This has inspired most researchers to focus on how express emotions and mental states exploiting anthropomorphic features, in many occasions relying just on faces. However emotions and mental states are not just presented through facial expression, but also by body postures and other features~\cite{Gelder2008}. Nonetheless in psychology has been a clear tendency to study the role of human face in emotion projection~\cite{Ekman2004},~\cite{kleinsmith2012affective} and mental states. This trend has been followed by robotics community, where anthropomorphic faces (e.g.,~\cite{Arras2012},~\cite{Breazeal2002}) and bodies (e.g.,~\cite{Canamero2010},~\cite{haering2011},~\cite{Destephe2013}) have been widely used to convey emotions.
Nevertheless, in many situations the presence of anthropomorphic elements would be out of place and not justified by the main robot's functionalities. Most of the current and future robotics platforms on the market will not require anthropomorphic faces or limbs. In some cases, like, for instance, in floor cleaning robots, anthropomorphic characteristics could even be detrimental to robot's task accomplishment.
This generates the necessity to study other mechanisms that could help to project emotions, which could give people an idea about the robots' state, and engage the user in long term relations.

The amount of works studying non-anthropomorphic features in robotics (e.g.,~\cite{Saerbeck2010,Lakatos2014,Sharma2013,Novika2015}) remains still small in comparison to those that exploit anthropomorphic features. Moreover, these works do not prescribe any specific range of values for the characteristics to be used to express the  implemented emotions. For example, Suk and collaborators~\cite{NAM2014}, in their study, give specific values for acceleration, curvature, but their connection to specific emotions is not given. Rather, they gave a relationship between their features and values in terms of valence and arousal. Another possibility is the use of professional human actors to study how they convey certain emotions. However, a direct mapping between humans and robots is not possible~\cite{Saerbeck2007,Canamero2010} due to robots' physical capabilities. In addition, the significance of the agreement obtained from the studies that used human actors is on discussion~\cite{Russell2003}. As a consequence determine precise values to project emotions in non-anthropomorphic platforms is required.

In order to get a better understanding of other features and values that could be used to convey specific emotions, this paper presents an experiment that was designed to identify specific values for some movement features that could be used to express the following emotions, selected among the ones suggested by Ekman as basic~\cite{Ekman2004}: happiness, anger, fear, and sadness. The considered features are: oscillation angle, linear and angular velocity, direction and orientation, identified as independent variables. The perceived emotions and their intensities were considered as dependent variables. It is to be observed that it was given the possibility to the subjects to provide their evaluation of emotional intensity for more than one emotion for each experience. The experiment took place in Politecnico di Milano during June and July of 2015, where students from diverse departments were asked to participate without any economical retribution. Krippendorff's alpha agreement~\cite{Krippendorff2007} ($\alpha$) was used to evaluate the agreement among the participants for each treatment. For each of the four emotions was generated a top 10 list based on alpha agreement and perceived intensity. The results suggest that fear is perceived when the robot is looking at the subjects while moving far from them fast. Sadness is associated to slow velocities with slow angular velocity and small oscillation angle. Anger is attributed to fast velocities, both angular and linear, small angle of oscillation and the robot facing the subjects while approaching them.

>>>>

Social environments are intricate spaces where people behave in ways that aim to their acceptance in social groups. Similarly, studies in social robotics have shown that robots acceptance increases when robots project a high social presence~\cite{Heerink08}. The straightest way to fulfill is through imitation of humans' characteristics, such us body form, behaviors and social characteristics. This has inspired most researchers to focus on how express emotions and mental states exploiting anthropomorphic features, in many occasions relying just on faces. However emotions and mental states are not just presented through facial expression, but also by body postures and other features~\cite{Gelder2008}. Nonetheless in psychology has been a clear tendency to study the role of human face in emotion projection~\cite{Ekman2004},~\cite{kleinsmith2012affective} and mental states. This trend has been followed by robotics community, where anthropomorphic faces (e.g.,~\cite{Arras2012},~\cite{Breazeal2002}) and bodies (e.g.,~\cite{Canamero2010},~\cite{haering2011},~\cite{Destephe2013}) have been widely used to convey emotions.
Nevertheless, in many situations the presence of anthropomorphic elements would be out of place and not justified by the main robot's functionalities. Most of the current and future robotics platforms on the market will not require anthropomorphic faces or limbs. In some cases, like, for instance, in floor cleaning robots, anthropomorphic characteristics could even be detrimental to robot's task accomplishment.
This generates the necessity to study other mechanisms that could help to project emotions, which could give people an idea about the robots' state, and engage the user in long term relations.

The amount of works studying non-anthropomorphic features in robotics (e.g.,~\cite{Saerbeck2010,Lakatos2014,Sharma2013,Novika2015}) remains still small in comparison to those that exploit anthropomorphic features. Moreover, these works do not prescribe any specific range of values for the characteristics to be used to express the  implemented emotions. For example, Suk and collaborators~\cite{NAM2014}, in their study, give specific values for acceleration, curvature, but their connection to specific emotions is not given. Rather, they gave a relationship between their features and values in terms of valence and arousal. Another possibility is the use of professional human actors to study how they convey certain emotions. However, a direct mapping between humans and robots is not possible~\cite{Saerbeck2007,Canamero2010} due to robots' physical capabilities. In addition, the significance of the agreement obtained from the studies that used human actors is still on discussion~\cite{Russell2003}. As a consequence is required the necessity to determine to project emotions in non-anthropomorphic platforms.

In order to get a better understanding of other features and values that could be used to convey specific emotions, this paper presents an experiment that was designed to identify specific values for some movement features that could be used to express the following emotions, selected among the ones suggested by Ekman as basic~\cite{Ekman2004}: happiness, anger, fear, and sadness. The considered features are: oscillation angle, linear and angular velocity, direction and orientation, identified as independent variables. The perceived emotions and their intensities were considered as dependent variables. It is to be observed that it was given the possibility to the subjects to provide their evaluation of emotional intensity for more than one emotion for each movement of the robot or treatment. The experiment took place in Politecnico di Milano during June and July of 2015, where students from diverse departments were asked to participate without any economical retribution. Krippendorff's alpha agreement~\cite{Krippendorff2007} ($\alpha$) was used to evaluate the agreement among the participants for each treatment. For each of the four emotions was generated a top 10 list based on alpha agreement and perceived intensity. The results suggest that fear is perceived when the robot is looking at the subjects while moving far from them fast. Sadness is associated to slow velocities with slow angular velocity and small oscillation angle. Anger is attributed to fast velocities, both angular and linear, small angle of oscillation and the robot facing the subjects while approaching them.

### 2017-04-23 JournalIJSR/Parts/Platform.tex
Last version of the article. This version includes most of reviewers´ comments, such as new references, explanation of parts that were not well explained before.

%%%%%%%%%%%%%%%%%%%%%%%platform_base
\subsection{Software}
To guarantee that the correct robot's velocity for each motor could be achieved, a PID controller was implemented. PID's set point is established by a higher level controller that could receive two different types of commands. The first commands are robot's velocities ($<V_x,V_y,\omega>$), given in robot's reference frame. The second command is a set point ($<x, Y, \theta>$) to be reached, given in the world reference frame. This world frame of reference is set every time the system is reset or boot. For example, if the robot finishes a trajectory and then it is reset, the new frame is going to be in the robot's current position.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%platform_base
\subsection{Software}
To guarantee that the correct robot's velocity for each motor could be achieved, a PID controller was implemented. PID's sets the velocity  that each motor must maintain when the platform receives a high level command. To maintain the velocity, the PID uses the velocity calculated using the odometry. Figure~\ref{fig:distribution} the general distribution of all the components, which include kinematics calculation, motors' velocity and platform velocity.

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{Images/version-2-program.png} 
	\caption{General distribution and communication scheme among the diverse software components running on the Arduino.}
	\label{fig:distribution}
\end{figure}

Two commands could be received,  the first command specifies robot's velocities ($<V_x,V_y,\omega>$), given in robot's reference frame. The second command is a set point ($<x, Y, \theta>$) to be reached, given in the world reference frame. This world frame of reference is set every time the system is reset or boot. For example, if the robot finishes a trajectory and then it is reset, the new frame is going to be in the robot's current position.

### 2017-04-23 JournalIJSR/Parts/Results.tex
Last version of the article. This version includes most of reviewers´ comments, such as new references, explanation of parts that were not well explained before.

For each table, mean, standard deviation, and median were calculated. It was not possible to use ANOVA test over the data because the assumption of normality is not satisfied in the collected data. This was checked using the Shapiro-Wilk Test.
Additionally, a contingency table for each emotion was generated in each treatment, as it is depicted in Table~\ref{table:table_contingency}, where the intensity for the rest of emotions is calculated as their mean of their intensity, including the option ``other''. Therefore a total of seven tables for each treatment were created. For all these tables, Krippendorff's alpha agreement~\cite{Krippendorff2007} ($\alpha$) was calculated, which is a reliability coefficient to measure the agreement among different participants. Unlike other coefficients (e.g., $Kappa$), $\alpha$ is a generalization of several known reliability indices, and it applies to any number of observers, any number of categories, any type of data, incomplete or missing data, large and small sample sizes.

This calculation was done using the R's package \textit{irr}. To improve contingency tables' interpretation, it was decided to just register emotions' alpha and intensity values greater than zero. The $\alpha$ for each treatment is shown in Figure~\ref{fig:alpha_experience}. As it could be observed, this table does not provide any information about the interpretation for each emotion. Therefore, it was decided to create top ten raking of treatments from the contingency tables of each emotion. This ranking is generated considering the following items in the respective order:(i) the mean intensity of the respective emotion, (ii) the alpha agreement of contingency table for the respective emotion, and (iii) the alpha agreement for the specific treatment. From the resulting tables~\ref{table:happy_top_ten},~\ref{table:Angry_top_ten},~\ref{table:Sad_top_ten}, and~\ref{table:Scared_top_ten}, it is possible to notice:

\begin{itemize}

	\item \textit{Fear} was the only emotion that had six over ten movements obtaining both general and specific alpha agreement over 0.41, which is the lower bound for moderate agreement~\cite{Viera2005}. Also, it was mostly selected (eight over ten) when $direction = \pi$ and $orientation = \pi$. The angle and the angular velocity attributed six over ten times were $0.087$ and $2$, respectively. So it seems that people perceive the movement as a fear expression when the robot is looking at them and moving far from them fast.

	\item \textit{Sadness} was mostly attributed to linear velocities of $0 mm/s$ (two over ten) and $200 mm/s$ (eight over ten), and angular velocities of $0 rad/s$ and $1 rad/s$, with higher predilection of the second (eight over ten). It seems that people attribute sadness to slow velocities with slow angular velocity and small oscillation angle. Regarding the other two features, there is not any defined pattern that could lead to make a generalization.

	\item \textit{Happiness} is attributed to different values of the independent variables. But it is the only emotion that is highly perceived when the linear velocity is $0$ with four over ten. However the oscillation angle for these four cases is equally divided between the values $0.174$ and $3.49$. %TODO BTW, is this angle in radians? It is a quite big difference...
So it seems that happiness is mainly attributed to fast angular velocities and big oscillations angles. Specific agreement among the other features is not present.

	\item \textit{Anger} is highly perceived with an angle of $0.087$ (seven over ten) and with linear velocities over $200 mm/s$. It seems that people attribute anger to fast velocities, both angular and linear, small angle of oscillation and the robot facing the persons while approaching them. 
\end{itemize} 
%alpha general

>>>>

For each table were calculated mean, standard deviation, and median. It was not possible to use ANOVA test over the data because the assumption of normality is not satisfied in the collected data. This was checked using the Shapiro-Wilk Test.
Additionally, a contingency table for each emotion was generated in each treatment, as it is depicted in Table~\ref{table:table_contingency}, where the intensity for the rest of emotions is calculated as their mean of their intensity, including the option ``other''. Therefore a total of seven tables for each treatment were created. For all these tables, Krippendorff's alpha agreement~\cite{Krippendorff2007} ($\alpha$) was calculated, which is a reliability coefficient to measure the agreement among different participants. Unlike other coefficients (e.g., $Kappa$), $\alpha$ is a generalization of several known reliability indices, and it applies to any number of observers, any number of categories, any type of data, incomplete or missing data, large and small sample sizes.

This calculation was done using the R's package \textit{irr}. To improve contingency tables' interpretation, it was decided to just register emotions' alpha and intensity values greater than zero. The $\alpha$ for each treatment is shown in Figure~\ref{fig:alpha_experience}. As it could be observed, this table does not provide any information about the interpretation for each emotion. Therefore, it was decided to create top ten raking of treatments from the contingency tables of each emotion. This ranking is generated considering the following items in the respective order:(i) the mean intensity of the respective emotion, (ii) the alpha agreement of contingency table for the respective emotion, and (iii) the alpha agreement for the specific treatment. Table~\ref{table:happy_top_ten} for Happiness, Table~\ref{table:Angry_top_ten} for Anger, Table~\ref{table:Sad_top_ten} for Sadness, and Table~\ref{table:Scared_top_ten} for Fear.
